{Gravitational wave sources in our Galactic backyard:}\\{Predictions for BHBH, BHNS and NSNS binaries detectable with LISA

### 2022-07-06 paper/tex/sections/abstract.tex
Updates from Overleaf

Future searches for gravitational waves from space will be sensitive to double compact objects (DCOs) in our Milky Way. We present new simulations of the populations of double black holes (BHBHs), black hole neutron stars (BHNSs) and double neutron stars (NSNSs) that will be detectable by the planned space-based gravitational wave detector LISA. 
%
For our estimates, we use an empirically-informed model of the metallicity dependent star formation history of the Milky Way. We populate it using an extensive suite of binary population-synthesis predictions for varying assumptions relating to mass transfer, common-envelope, supernova kicks, remnant masses and wind mass loss physics.

For a 4(10)-year LISA mission, we predict between \rangeFourYear{}(\rangeTenYear{}) detections over these variations, out of which 6-154(9-238) are BHBHs, 2-198(3-289) are BHNSs and 3-35(4-57) are NSNSs.
%
\edit1{We expect that about 50\%(60\%) can be  distinguished from WDWD sources, based on their mass or eccentricity and localisation.  Specifically, for about 10\%(15\%) we expect to be able to determine chirp masses better than 10\%. For 13\%(13\%) we expect sky-localisations better than 1 degree. % 
We discuss the impact of physics variations and prospects for probing the NS-BH mass gap.}  We further discuss the possibility of multi-messenger observations of pulsar populations with the Square Kilometre Array (SKA) and assess the benefits of extending the LISA mission.

>>>>

Future searches for gravitational waves from space will be sensitive to double compact objects (DCOs) in our Milky Way. We present new simulations of the populations of double black holes (BHBHs), black hole neutron stars (BHNSs) and double neutron stars (NSNSs) that will be detectable by the planned space-based gravitational wave detector LISA. 
%
For our estimates, we use an empirically-informed model of the metallicity dependent star formation history of the Milky Way. We populate it using an extensive suite of binary population-synthesis predictions for varying assumptions relating to mass transfer, common-envelope, supernova kicks, remnant masses and wind mass loss physics. \edit1{Each model in the grid has been recently shown to be consistent with current constraints on the inferred overall GW rates from ground-based detectors.}

For a 4(10)-year LISA mission, we predict between \rangeFourYear{}(\rangeTenYear{}) detections over these variations, out of which 6-154(9-238) are BHBHs, 2-198(3-289) are BHNSs and 3-35(4-57) are NSNSs.
%
\edit1{We expect that about 50\%(60\%) can be distinguished from WDWD sources, based on their mass or eccentricity and localisation. Specifically, for about 10\%(15\%) we expect to be able to determine chirp masses better than 10\%. For 13\%(13\%) we expect sky-localisations better than 1 degree.} We discuss how the variations in the physics assumptions alter the distribution of properties of the detectable systems, even when the detection rates are unchanged. We further discuss the possibility of multi-messenger observations of pulsar populations with the Square Kilometre Array (SKA) and assess the benefits of extending the LISA mission.

### 2022-07-06 paper/tex/sections/introduction.tex
Updates from Overleaf

BHBHs in the Milky Way present a greater observational challenge. To date, no BH has been observed in a BHBH binary in the Milky Way, and so LISA could provide the first detection of a Galactic BHBH. The only confirmed BHs in our galaxy have been discovered as components of X-ray binaries with companion stars \citep[e.g.][]{Bolton+1972,Webster+1972}. These detections have observed BHs with masses mainly constrained between $5$ and $10 \unit{M_\odot}$ \citep{Corral-Santana+2016}, a stark contrast to the more massive BHs observed with LIGO/Virgo that tend to contain at least one BH with a mass greater than $10 \unit{M_{\odot}}$ \citep{Abbott+2020_GWTC2}. These observations of X-ray binaries suggest the presence of a lower mass gap (from $2$-$5 \unit{M_{\odot}}$) in which there are no strong candidates for either black holes or neutron stars \citep{Ozel+2010,Farr+2011} but the gap's existence remains an open question \citep[e.g.][]{Kreidberg+2012, Mandel+2020}. Recently there has also been increased discussion over the maximum BH mass in our galaxy, with the claims of a $70 \unit{M_{\odot}}$ BH \citep{Liu+2019} which has subsequently been challenged (\citealp{El-Badry+2020, Abdul-Masih+2020, Shenar+2020,Eldridge+2020}, see also \citealp{Liu+2020}) and revised measurements of the mass of Cygnus X-1 \citep{Miller-Jones+2021}. A sample of BHBHs detected with LISA could possibly help to constrain the BH mass distribution.

@@

Earlier work on BHBHs, BHNSs and NSNSs in LISA has used a variety of population synthesis codes, Milky Way models and LISA specifications, resulting in a wide range of predictions \citep{Nelemans+2001,Belczynski+2010,Liu+2014,Lamberts+2019,Lau+2020,Breivik+2020,Sesana+2020, Shao+2021}. We build upon previous efforts but with several important improvements. We explore the effects of varying binary physics assumptions by repeating our analysis for \nModels{} different models and comparing the effects on the detection rate and distributions of source parameters \citep{Broekgaarden+2021,Broekgaarden+2021b}. We use a model for the Milky Way that accounts for the chemical enrichment history and is calibrated on the APOGEE survey \citep{Majewski+2017,Frankel+2018}, whereas most others did not consider the effect of metallicity in detail (see however \citealp{Lamberts+2019, Sesana+2020}). We provide a detailed treatment of the eccentricity of detectable sources, both for the inspiral evolution as well as gravitational wave signal during the LISA mission. Moreover, the grid of binary population synthesis simulations that we use is the most extensive of its kind to date and makes use of the adaptive sampling algorithm STROOPWAFEL \citep{Broekgaarden+2019, Broekgaarden+2021}. Overall over 2 billion massive binaries were simulated to produce the DCO populations used in this work. We find that this large number of simulations is important to reduce the sampling noise even when using adaptive importance sampling.

@@

Our paper is structured as follows. In Section~\ref{sec:method}, we describe the methods for synthesising a population of binaries, the variations of physical assumptions that we consider, how we simulate the Milky Way distribution of DCOs and our methods for calculating a detection rate for LISA. We present the main results for our fiducial model in Section~\ref{sec:results}, before exploring the variations in the detectable population when changing our physical assumptions in Section~\ref{sec:variations}. \edit1{In Section~\ref{sec:discussion} we discuss these how and whether these sources can be distinguished from the WDWD population and whether the subpopulations of NSNS, BHNS and BHBH can be distinguished from each other. We further discuss prospects for identifying EM counterparts, the role of different formation channels and overall caveats in the modelling approach. }In Section~\ref{sec:compare_studies}, we compare and contrast our methods and findings to previous work and finish with our conclusions in Section~\ref{sec:conclusion}.

>>>>

BHBHs in the Milky Way present a greater observational challenge. To date, no BH has been observed in a BHBH binary in the Milky Way, and so LISA could provide the first detection of a Galactic BHBH. The only confirmed BHs in our galaxy have been discovered as components of X-ray binaries with companion stars \citep[e.g.][]{Bolton+1972,Webster+1972}. These detections have observed BHs with masses mainly constrained between $5$ and $10 \unit{M_\odot}$ \citep{Corral-Santana+2016}, a stark contrast to the more massive BHs observed with LIGO/Virgo that tend to contain at least one BH with a mass greater than $10 \unit{M_{\odot}}$ \citep{Abbott+2020_GWTC2}. These observations of X-ray binaries suggest the presence of a lower mass gap (from $2$-$5 \unit{M_{\odot}}$) in which there are no strong candidates for either black holes or neutron stars \citep{Ozel+2010,Farr+2011} but the gap's existence remains an open question \edit1{\citep[e.g.][]{Kreidberg+2012, Mandel+2020, Jayasinghe+2021}}. Recently there has also been increased discussion over the maximum BH mass in our galaxy, with the claims of a $70 \unit{M_{\odot}}$ BH \citep{Liu+2019} which has subsequently been challenged (\citealp{El-Badry+2020, Abdul-Masih+2020, Shenar+2020,Eldridge+2020}, see also \citealp{Liu+2020}) and revised measurements of the mass of Cygnus X-1 \citep{Miller-Jones+2021}. A sample of BHBHs detected with LISA could possibly help to constrain the BH mass distribution.

@@

Earlier work on BHBHs, BHNSs and NSNSs in LISA has used a variety of population synthesis codes, Milky Way models and LISA specifications, resulting in a wide range of predictions \citep{Nelemans+2001,Belczynski+2010,Liu+2014,Lamberts+2019,Lau+2020,Breivik+2020,Sesana+2020, Shao+2021}. We build upon previous efforts but with several important improvements. We explore the effects of varying binary physics assumptions by repeating our analysis for \nModels{} different models\edit1{, which have each been shown to be consistent with constraints on the inferred overall GW rates from ground-based detectors,} and comparing the effects on the detection rate and distributions of source parameters \citep{Broekgaarden+2021,Broekgaarden+2021b}. We use a model for the Milky Way that accounts for the chemical enrichment history and is calibrated on the APOGEE survey \citep{Majewski+2017,Frankel+2018}, whereas most others did not consider the effect of metallicity in detail (see however \citealp{Lamberts+2019, Sesana+2020}). We provide a detailed treatment of the eccentricity of detectable sources, both for the inspiral evolution as well as gravitational wave signal during the LISA mission. Moreover, the grid of binary population synthesis simulations that we use is the most extensive of its kind to date and makes use of the adaptive sampling algorithm STROOPWAFEL \citep{Broekgaarden+2019, Broekgaarden+2021}. Overall over 2 billion massive binaries were simulated to produce the DCO populations used in this work. We find that this large number of simulations is important to reduce the sampling noise even when using adaptive importance sampling.

@@

Our paper is structured as follows. In Section~\ref{sec:method}, we describe the methods for synthesising a population of binaries, the variations of physical assumptions that we consider, how we simulate the Milky Way distribution of DCOs and our methods for calculating a detection rate for LISA. We present the main results for our fiducial model in Section~\ref{sec:results}, before exploring the variations in the detectable population when changing our physical assumptions in Section~\ref{sec:variations}. \edit1{In Section~\ref{sec:discussion} we discuss these how and whether these sources can be distinguished from the WDWD population and whether the subpopulations of NSNS, BHNS and BHBH can be distinguished from one other. We further discuss prospects for identifying EM counterparts, the role of different formation channels and overall caveats in the modelling approach.} In Section~\ref{sec:compare_studies}, we compare and contrast our methods and findings to previous work and finish with our conclusions in Section~\ref{sec:conclusion}.

### 2022-07-06 paper/tex/sections/methods.tex
Updates from Overleaf

We use the grid of \nModels{} binary population synthesis simulations recently presented in \citet{Broekgaarden+2021,Broekgaarden+2021b}. This grid of simulations is synthesised using the rapid population synthesis code \href{https://compas.science}{COMPAS} \citep{Stevenson+2017,Vigna-Gomez+2018,Stevenson+2019,Broekgaarden+2019}. COMPAS follows the approach of the population synthesis code BSE \citep{Hurley+2000,Hurley+2002} and uses fitting formula and rapid algorithms to efficiently predict the final fate of binary systems. The code is open source and documented in the papers listed above, the online documentation\footnote{\url{https://compas.science}} and in the methods paper \citep{COMPAS:2021methodsPaper}. We summarise the main assumptions and settings relevant for this work in Appendix~\ref{app:pop_synth}.

@@

Most previous studies that predict a detection rate for LISA place binaries in the Milky Way independently of their age or evolution. We improve upon this as the first study to use an empirically-informed analytical model of the Milky Way that takes into account the galaxy's enrichment history by applying the metallicity-radius-time relation from \citet{Frankel+2018}. Those authors developed this relation in order to measure the global efficiency of radial migration in the Milky Way and calibrated it using a sample of red clump stars measured with APOGEE \citep{Majewski+2017}.

@@

Our model for the Milky Way accounts for the low-\achem\footnote{Nomenclature used to describe the enhancement of $\alpha$ elements compared to iron in stellar atmospheres}~disc, high-\achem~disc and a central component approximating a bar/bulge. The low- and high-\achem~discs are often also referred to as the thin and thick discs because the stellar vertical distribution is better fit by a double exponential rather than a single one. However, this doesn't allow one to assign a star to either the thin or thick disk purely based on its height above the Galactic plane. Therefore, we instead use the chemical definition of the two disks (applying the \achem~nomenclature) as there is a clear bimodal distribution in the chemical plane, allowing stars to be assigned to each of the disc components based on their chemical abundances. For each of the three components, we use a separate star formation history and spatial distribution, which we combine into a single model, weighting each component by its present-day stellar mass. \citet{Licquia+2015} gives that the stellar mass of the bulge is $0.9 \times 10^{10} \unit{M_{\odot}}$ and the stellar mass of the disc is $5.2 \times 10^{10} \unit{M_\odot}$, which we split equally between the low- and high-\achem~discs \citep[e.g.,][]{Snaith+2014}.

@@

The star formation history of the bulge/bar of the Milky Way has many uncertainties due to the (1) sizeable age measurement uncertainties at large ages in observational studies, (2) complex selection processes affecting the observed age distributions, and (3) formation mechanisms that are still under debate. However, the central bar (which we assume here to dominate) was shown to contain stars with an extended age range, with most observed stars between $6$ and $12 \unit{Gyr}$ with a younger tail of ages that could come from the subsequent secular growth of the Galactic bar \citep[e.g.,][]{Bovy+2019}. To model the bar/bulge's age distribution more realistically than in previous studies (which assume an old bulge coming from a single starburst), we choose to adopt a more extended star formation history using a $\beta(2,3)$ distribution, shifted and scaled such that stars are only formed in the range $[6, 12] \unit{Gyr}$. We show these distributions in the top left panel of Fig.~\ref{fig:galaxy_schematic}.

@@

This scale length accounts for the inside-out growth of the low-\achem~disc and hence is age dependent. We assume $R_d = (1 / 0.43) \unit{kpc}$ for the high-\achem~disc \citep[][Table~1]{Bovy+2016} and $R_d = 1.5 \unit{kpc}$ for the bulge/bar component \citep{Bovy+2019}. \edit1{Note that in this way we approximate the bulge/bar component as being axi-symmetric, which is sufficient for our purposes.} We show the combination of these distributions in the second panel on the left in Fig.~\ref{fig:galaxy_schematic}.

>>>>

We use the grid of \nModels{} binary population synthesis simulations recently presented in \citet{Broekgaarden+2021,Broekgaarden+2021b}. This grid of simulations is synthesised using the rapid population synthesis code \href{https://compas.science}{COMPAS} \citep{Stevenson+2017,Vigna-Gomez+2018,Stevenson+2019,Broekgaarden+2019}. COMPAS follows the approach of the population synthesis code BSE \citep{Hurley+2000,Hurley+2002} and uses fitting formula and rapid algorithms to efficiently predict the final fate of binary systems. The code is open source and documented in the papers listed above, the online documentation\footnote{\url{https://compas.science}} and in the methods paper \citep{COMPAS:2021methodsPaper}. \edit1{The model we refer to as our fiducial model allows for BHs with masses in the heavily debated mass gap, but we also consider \nModels{} physics variations. \citet{Broekgaarden+2021b} showed that all of these physics variations are still consistent with current constraints on the overall inferred GW rates, at least when considering large variations of the assumed metallicity specific cosmic star formation rate \citep{Abbott+2021_GWTC3}}. We summarise the main assumptions and settings relevant for this work in Appendix~\ref{app:pop_synth}.

@@

Most previous studies that predict a detection rate for LISA place binaries in the Milky Way independently of their age or evolution. We improve upon this as the first study to use an empirically-informed analytical model of the Milky Way that takes into account the galaxy's enrichment history by applying the metallicity-radius-time relation from \citet{Frankel+2018}. Those authors developed this relation in order to measure the global efficiency of radial migration in the Milky Way and calibrated it using a sample of red clump stars measured with APOGEE \citep{Majewski+2017}. \edit1{We assess the impact of using this improved Milky Way model in Appendix~\ref{app:mw_changes} and the effect of Galactic models on LISA predictions has been investigated more generally in \citet{Storck+2022}}.

@@

Our model for the Milky Way accounts for the low-\achem\footnote{Nomenclature used to describe the enhancement of $\alpha$ elements compared to iron in stellar atmospheres}~disc, high-\achem~disc and a central component approximating a bulge/bar. The low- and high-\achem~discs are often also referred to as the thin and thick discs because the stellar vertical distribution is better fit by a double exponential rather than a single one. However, this doesn't allow one to assign a star to either the thin or thick disk purely based on its height above the Galactic plane. Therefore, we instead use the chemical definition of the two disks (applying the \achem~nomenclature) as there is a clear bimodal distribution in the chemical plane, allowing stars to be assigned to each of the disc components based on their chemical abundances. For each of the three components, we use a separate star formation history and spatial distribution, which we combine into a single model, weighting each component by its present-day stellar mass. \citet{Licquia+2015} gives that the stellar mass of the bulge is $0.9 \times 10^{10} \unit{M_{\odot}}$ and the stellar mass of the disc is $5.2 \times 10^{10} \unit{M_\odot}$, which we split equally between the low- and high-\achem~discs \citep[e.g.,][]{Snaith+2014}.

@@

The star formation history of the bulge/bar of the Milky Way has many uncertainties due to the (1) sizeable age measurement uncertainties at large ages in observational studies, (2) complex selection processes affecting the observed age distributions, and (3) formation mechanisms that are still under debate. However, the central bar (which we assume to dominate here) was shown to contain stars with an extended age range, with most observed stars between $6$ and $12 \unit{Gyr}$ with a younger tail of ages that could come from the subsequent secular growth of the Galactic bar \citep[e.g.,][]{Bovy+2019}. To model the bulge/bar's age distribution more realistically than in previous studies (which assume an old bulge coming from a single starburst), we choose to adopt a more extended star formation history using a $\beta(2,3)$ distribution, shifted and scaled such that stars are only formed in the range $[6, 12] \unit{Gyr}$. We show these distributions in the top left panel of Fig.~\ref{fig:galaxy_schematic}.

@@

This scale length accounts for the inside-out growth of the low-\achem~disc and hence is age dependent. We assume $R_d = (1 / 0.43) \unit{kpc}$ for the high-\achem~disc \citep[][Table~1]{Bovy+2016} and $R_d = 1.5 \unit{kpc}$ for the bulge/bar component \citep{Bovy+2019}. \edit1{Note that in this way we have approximated the bulge/bar component as being axi-symmetric, which is sufficient for our purposes.} We show the combination of these distributions in the second panel on the left in Fig.~\ref{fig:galaxy_schematic}.

### 2022-07-06 paper/tex/sections/results.tex
Updates from Overleaf

\paragraph{Orbital Frequency} The orbital frequency distributions for BHBHs, BHNSs and NSNSs (Fig.~\ref{fig:fiducial_pdf_distributions}a) peak at progressively increasing frequencies. As mentioned in Sec.~\ref{sec:dcos_on_sc}, this is because a higher mass DCO at the same distance and eccentricity requires a lower frequency to produce the same signal-to-noise ratio and thus be detected. The distributions appear nearly symmetric, but closer inspections shows that the left hand side is more populated, which can be seen most clearly in the curve for the BHBHs. This is due to the contribution of highly eccentric binaries, which are most abundant in the BHBH population. These systems are still detectable by LISA, despite their low orbital frequency, as the high eccentricity means that the majority of the GW signal is emitted at higher harmonics, where LISA is more sensitive.

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{fig3_detectable_properties_4yr.pdf}
    \caption{Properties of detectable systems for a 4-year LISA mission in our fiducial model. Each panel shows a kernel density estimator for a single property, coloured by DCO type. Each curve has been individually normalised. The shaded areas show the 1- and 2-$\sigma$ sampling uncertainties (obtained via bootstrapping). The dotted lines in panel b show the individual primary and secondary mass distributions. The dotted line in panel e shows the star formation history we have assumed in our Milky Way model. See Sec.~\ref{sec:fiducial_distributions} for a discussion. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig3_detectable_properties_4yr.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:fiducial_pdf_distributions}
\end{figure*}

\paragraph{Black Hole Mass}
In Fig.~\ref{fig:fiducial_pdf_distributions}b, we show the distribution masses of individual black holes that are part of BHNS and BHBH systems. 
The distribution for BHBH shows a bimodality. This results from the two contributions of the more and less massive BHs in BHBH systems, which peak at around $8 \unit{M_{\odot}}$ and $3.5 \unit{M_{\odot}}$ respectively as shown by the dotted lines (see also our discussion of the mass ratios below).

For both the BHBHs and BHNSs, we see that the black hole mass distribution favours low masses. About $90\%$ of BHs have masses below $11 \unit{M_{\odot}}$, in our fiducial simulations shown here. This is in stark contrast with observations from ground-based GW detectors, where heavy BHs with masses of about $30\unit{M_{\odot}}$ and higher have been common. There are two main reasons for this discrepancy. First, the population of BHs in the Milky Way (and, in particular, those detectable by LISA due to their recent formation times, see Fig.~\ref{fig:fiducial_pdf_distributions}e) primarily come from progenitors that formed from high metallicity gas according to our simulations. Stellar winds are stronger at high metallicity leading to increased mass loss. This affects the mass of the most massive black holes that can be formed \citep{Belczynski+2010}. Secondly, ground-based detectors are strongly biased towards high mass systems, since they can be detected out to larger distances and thus a greater volume is probed. In contrast, LISA has no such bias and is, in principle, sensitive to inspiraling BHBHs throughout the entire Milky Way regardless of their mass, as long as we catch them when they are emitting in the LISA band. For this reason, LISA is more likely to detect the more common lower mass BHBHs.

@@

In contrast, only $\BHBHqAbovePointEight{}$ of detectable BHBHs are formed with $q > 0.8$ and the distribution peaks around $q = 0.4$. We find that the strong stellar winds in our (typically high-metallicity) progenitors are the reason behind this.

BHBHs with unequal masses typically come from progenitors that also had more extreme mass ratios at birth ($90$ and $30\unit{M_{\odot}}$ are typical for the progenitors of detectable BHBH systems in our simulations). The primary in such systems experiences strong mass loss by winds before filling its Roche lobe. This mostly happens during its early hydrogen-shell burning phase. The wind mainly reduces the mass of the envelope, but does not have a very strong effect on the core. By the time the primary fills its Roche lobe, it has become less massive and the mass ratio is closer to one. This favours stable mass transfer. The massive core of the primary star typically becomes the more massive BH. Accretion on the secondary star is limited and the secondary eventually provides the less massive black hole.

@@

When comparing the distribution of formation times for the three different DCO types we see that NSNSs are most strongly concentrated at recent times, followed by BHBHs and then BHNSs. To understand this it is helpful to consider how the inspiral time scales with various parameters \citep{Peters+1964}
\begin{equation}
    t_{\rm inspiral} \propto \frac{a^4}{(m_1 + m_2)^{3}} \cdot \frac{q}{(1+q)^2} \cdot (1 - e^2)^{-7/2}.
\end{equation}
The inspiral time depends most strongly on the separation at DCO formation, $a$, and this is where the three types also differ most strongly (see Fig.~\ref{fig:dco_formation_properties}). The detectable NSNS systems have the tightest orbits at DCO formation. The median of the distribution of separations at DCO formation, $\langle a_{\rm DCO} \rangle_{\rm med}$, relate as 8:3:1 for detectable BHBH:BHNS:NSNS in our simulations. This results in increase of the inspiral time by a factor of about 4000:80:1. The total masses affect the inspiral time to the third power and this where the heavier BHBH systems are favoured. The median total masses differ by ratios of 6:4:1 for detectable BHBH:BHNS:NSNS in our simulations, impacting the inspiral times such that they are a factor of 200:60:1 shorter, partially counteracting the effect of the separations. The term depending on the mass ratio $q$ only varies by about 30\% for the mass ratio ranges considered here and so is not of interest. The eccentricity term is not of importance for mildly eccentric systems, $f(e_{\rm DCO} \le 0.3)  \le 1.4$ but of large importance for the very eccentric  $f(e_{\rm DCO} \ge 0.9) \ge 300$. The fraction of highly eccentric systems with $e_{\rm DCO}>0.9$ is 33\%, 16\%, and 8\% of for BHBH, BHNS and NSNS respectively, see also Fig.~\ref{fig:dco_formation_properties}.

@@

\subsection{Formation channels}\label{sec:progenitors_and_formation}
In our fiducial model, approximately two thirds of detectable BHBHs are formed through the `only stable mass transfer' channel, whilst the remaining third are primarily formed through the `classic' CE channel. Detectable BHNSs follow an inverse pattern, such that around two thirds are formed through the classic channel and the rest are mainly formed through only stable mass transfer.

>>>>

\paragraph{Orbital Frequency} The orbital frequency distributions for BHBHs, BHNSs and NSNSs (Fig.~\ref{fig:fiducial_pdf_distributions}a) peak at progressively increasing frequencies. As mentioned in Sec.~\ref{sec:dcos_on_sc}, this is because a higher mass DCO at the same distance and eccentricity requires a lower frequency to produce the same signal-to-noise ratio and thus be detected. The distributions appear nearly symmetric, \edit1{but the left tail of the distribution is more populated, especially} for the BHBHs. This is due to the contribution of highly eccentric binaries, which are most abundant in the BHBH population. These systems are still detectable by LISA, despite their low orbital frequency, as the high eccentricity means that the majority of the GW signal is emitted at higher harmonics, where LISA is more sensitive.

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{fig3_detectable_properties_4yr.pdf}
    \caption{Properties of detectable systems for a 4-year LISA mission in our fiducial model. Each panel shows a kernel density estimator for a single property, coloured by DCO type. Each curve has been individually normalised. The shaded areas show the 1- and 2-$\sigma$ sampling uncertainties (obtained via bootstrapping). The dotted line in panel e shows the star formation history we have assumed in our Milky Way model. See Sec.~\ref{sec:fiducial_distributions} for a discussion. \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figures/fig3_detectable_properties_4yr.pdf}{\faFileImage} \href{https://github.com/TomWagg/detecting-DCOs-in-LISA/blob/main/paper/figure_notebooks/fiducial.ipynb}{\faBook}.}
    \label{fig:fiducial_pdf_distributions}
\end{figure*}

\paragraph{Black Hole Mass}
In Fig.~\ref{fig:fiducial_pdf_distributions}b, we show the distribution of \edit1{primary} black hole masses for BHNS and BHBH systems. \edit1{The BHNS distribution shows a bimodality. This is due to the bimodal mass ratio distribution (see Fig.~\ref{fig:fiducial_pdf_distributions}c) since most NS components have approximately the same mass.}

For both the BHBHs and BHNSs, we see that the black hole mass distribution favours low masses. About $90\%$ of BHs have masses below $12 \unit{M_{\odot}}$, in our fiducial simulations shown here. This is in contrast with observations from ground-based GW detectors, where heavy BHs with masses of about $30\unit{M_{\odot}}$ and higher have been common. There are two main reasons for this discrepancy. First, the population of BHs in the Milky Way (and, in particular, those detectable by LISA due to their recent formation times, see Fig.~\ref{fig:fiducial_pdf_distributions}e) primarily come from progenitors that formed from high metallicity gas according to our simulations. Stellar winds are stronger at high metallicity leading to increased mass loss. This affects the mass of the most massive black holes that can be formed \citep{Belczynski+2010}. Secondly, ground-based detectors are strongly biased towards high mass systems, since they can be detected out to larger distances and thus a greater volume is probed. In contrast, LISA has no such bias and is, in principle, sensitive to inspiraling DCOs throughout the entire Milky Way regardless of their mass, as long as we catch them when they are emitting in the LISA band. For this reason, LISA is more likely to detect the more common lower mass DCOs.

@@

In contrast, only $\BHBHqAbovePointEight{}$ of detectable BHBHs are formed with $q > 0.8$ and the distribution peaks around $q = 0.4$. We find that the strong stellar winds in our (typically high-metallicity) progenitors are the reason behind this\edit1{, though we still would not expect a significant number of detections at higher mass ratios since the intrinsic merging BHBH population tends to have less equal mass ratios \citep[e.g.][Fig.~8]{vanSon+2021}}.

BHBHs with unequal masses typically come from progenitors that also had more extreme mass ratios at birth ($90$ and $30\unit{M_{\odot}}$ are typical for the progenitors of detectable BHBH systems in our simulations). The primary in such systems experiences strong mass loss by winds before filling its Roche lobe. This mostly happens during its early hydrogen-shell burning phase. The wind mainly reduces the mass of the envelope, but does not have a very strong effect on the core. By the time the primary fills its Roche lobe, it has become less massive and the mass ratio is closer to one. This favours stable mass transfer\edit1{ and most detectable BHBHs form through stable mass transfer (see Section.~\ref{sec:progenitors_and_formation})}. The massive core of the primary star typically becomes the more massive BH. Accretion on the secondary star is limited and the secondary eventually provides the less massive black hole.

@@

When comparing the distribution of formation times for the three different DCO types we see that NSNSs are most strongly concentrated at recent times, followed by BHBHs and then BHNSs. To understand this it is helpful to consider how the inspiral time scales with various parameters \citep{Peters+1964}
\begin{equation}
    t_{\rm inspiral} \propto \frac{a^4}{(m_1 + m_2)^{3}} \cdot \frac{q}{(1+q)^2} \cdot (1 - e^2)^{-7/2}.
\end{equation}
The inspiral time depends most strongly on the separation at DCO formation, $a$, and this is where the three types also differ most strongly (see Fig.~\ref{fig:dco_formation_properties}). The detectable NSNS systems have the tightest orbits at DCO formation. The median of the distribution of separations at DCO formation, $\langle a_{\rm DCO} \rangle_{\rm med}$, relate as 8:3:1 for detectable BHBH:BHNS:NSNS in our simulations. This results in increase of the inspiral time by a factor of about 4000:80:1.

\edit1{The distribution of separations at DCO formation is also the reason that the time since formation for NSNSs shows a bimodality. For NSNSs, the separation at DCO formation peaks around $2 \times 10^{-2} \unit{AU}$ but the distribution has a second population that extends to much lower separations relative to the the BHNS and BHBH distributions. It is this population of shorter separations that corresponds to the NSNSs that were formed more recently (within the last 100 Myr).}

The total masses affect the inspiral time to the third power and this where the heavier BHBH systems are favoured. The median total masses differ by ratios of 6:4:1 for detectable BHBH:BHNS:NSNS in our simulations, impacting the inspiral times such that they are a factor of 200:60:1 shorter, partially counteracting the effect of the separations. The term depending on the mass ratio $q$ only varies by about 30\% for the mass ratio ranges considered here and so is not of interest. The eccentricity term is not of importance for mildly eccentric systems, $f(e_{\rm DCO} \le 0.3)  \le 1.4$ but of large importance for the very eccentric  $f(e_{\rm DCO} \ge 0.9) \ge 300$. The fraction of highly eccentric systems with $e_{\rm DCO}>0.9$ is 33\%, 16\%, and 8\% of for BHBH, BHNS and NSNS respectively, see also Fig.~\ref{fig:dco_formation_properties}.

@@

\subsection{Formation channels}\label{sec:progenitors_and_formation}
In our fiducial model, approximately two thirds of detectable BHBHs are formed through the `only stable mass transfer' channel, whilst the remaining third are primarily formed through the `classic' CE channel \edit1{, in which the first phase of mass transfer is stable whilst the second is unstable and results in a common-envelope event}. Detectable BHNSs follow an inverse pattern, such that around two thirds are formed through the classic channel and the rest are mainly formed through only stable mass transfer.
